---
output: html_document
runtime: shiny
---


```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# VISUALIZING IOWA EDUCATION

This collection of maps explores some variables releated to educational resource allocation and measures used to evaluate student success. 

```{r, echo=FALSE}
rm(list=ls())
library(readr)
library(choroplethr)
library(choroplethrMaps)
library(dplyr)
library(ggplot2) 
library(plotly)
library(tmap)
library(sf)
library(tmaptools)
library(gganimate)
library(cowplot)
library(gridExtra)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Importing in geojoson file 
my_map <- st_read("~/Documents/research/primate_map/project_materials/sci_comm/shapefiles/rural_variables.geojson", quiet=TRUE)
my_data <- data.frame(my_map)

# Convert size variable to numeric and remove missing values
my_map$cleaned_rural_density <- as.numeric(my_map$cleaned_rural_density)
my_map$cleaned_rural_mean_sci <- as.numeric(my_map$cleaned_rural_mean_sci)
my_map$cleaned_rural_mean_math <- as.numeric(my_map$cleaned_rural_mean_math)
my_map$cleaned_rural_CountyGraduateRate4yr <- as.numeric(my_map$cleaned_rural_CountyGraduateRate4yr)
my_map$cleaned_rural_CountyGraduateRate5yr <- as.numeric(my_map$cleaned_rural_CountyGraduateRate5yr)
my_map$cleaned_rural_CountySpendingPerStudent <- as.numeric(my_map$cleaned_rural_CountySpendingPerStudent)
my_map$cleaned_rural_Graduate.or.professional.degree.. <- as.numeric(my_map$cleaned_rural_Graduate.or.professional.degree..)
my_map$cleaned_rural_Bachelor.s.degree.. <- as.numeric(my_map$cleaned_rural_Bachelor.s.degree..)

# Palettes of interest 
# palette = c("#a6bddb", "#3690c0", "#034e7b"), 
# palette = c("#8c510a", "#d8b365", "#7fbf7b"),


```


# ACT Math & Science Score Averages 
The below two maps provide the average ACT Math and Science scores for each county depicted as a proportional symbols. They are presented on a choropleth map showing the average amount of money spent per student. 

```{r}

library(shiny)
library(tmap)

# Define UI for the app
ui <- fluidPage(
  # Create two tabs for the maps
  navbarPage(
    "Maps",
    tabPanel("Science Scores", 
             tmapOutput("map_science")),
    tabPanel("Math Scores",
             tmapOutput("map_math"))
  )
)

# Define server logic
server <- function(input, output) {
  # Render the science scores map
  output$map_science <- renderTmap({
    # Set quantiles 
    my_map$binned_science <- as.numeric(my_map$cleaned_rural_mean_sci)
    my_map$binned_science <- cut(my_map$cleaned_rural_mean_sci,
                                 breaks = c(19.01000, 20.13391, 20.64000, 21.16250, 22.63571),
                                 labels = c(1, 2, 3, 4))
    my_map$binned_science <- as.numeric(my_map$binned_science)
    
    # Create the map
    tm_shape(my_map) +
      tm_fill("cleaned_rural_CountySpendingPerStudent", style = "quantile", n = 5, palette = "Greys", title = "Avg. Spending per Student") +
      tm_bubbles(size = "binned_science",
                 col = "#034e7b", 
                 alpha = .8, 
                 title.size ="Avg. ACT Science Score", 
                 legend.size.is.portrait=TRUE,
                 sizes.legend.labels=c("<= 20.1","<= 20.6","<= 21.2","<= 22.6"), 
                 scale = 1.75) +
      tm_legend(outside = TRUE, 
                outside.position = "left", 
                bg.color = "transparent", 
                frame = FALSE)
  })
  
  # Render the math scores map
  output$map_math <- renderTmap({
    # Set quantiles 
    my_map$binned_math <- as.numeric(my_map$cleaned_rural_mean_math)
    my_map$binned_math <- cut(my_map$cleaned_rural_mean_math,
                                 breaks = c(19.66250, 21.33818, 21.68000, 22.08706, 23.12000),
                                 labels = c(1, 2, 3, 4))
    my_map$binned_math <- as.numeric(my_map$binned_math)
    
    # Create the map
    tm_shape(my_map) +
      tm_fill("cleaned_rural_CountySpendingPerStudent", style = "quantile", n = 5, palette = "Greys", title = "Avg. Spending per Student") +
      tm_bubbles(size = "binned_math",
                 col = "#034e7b",
                 alpha = .8, 
                 title.size ="Avg. ACT Math Score", 
                 legend.size.is.portrait=TRUE,
                 sizes.legend.labels=c("<= 21.3","<= 21.7","<= 22.1","<= 23.1"), 
                 scale = 1.75) +
      tm_legend(outside = TRUE, 
                outside.position = "left", 
                bg.color = "transparent", 
                frame = FALSE)
  })
}

# Run the app
shinyApp(ui = ui, server = server)


```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# Determine quantiles 
# quantile(my_map$cleaned_rural_mean_sci, na.rm = TRUE)

# Set quantiles 
# my_map$binned_science <- as.numeric(my_map$cleaned_rural_mean_sci)
# my_map$binned_science <- cut(my_map$cleaned_rural_mean_sci,
                             # breaks = c(19.01000, 20.13391, 20.64000, 21.16250, 22.63571),
                             # labels = c(1, 2, 3, 4))
# my_map$binned_science <- as.numeric(my_map$binned_science)


# tm_shape(my_map) +
#   tm_fill("cleaned_rural_CountySpendingPerStudent", style = "quantile", n = 5, palette = "Greys", title = "Avg. Spending per Student") +
#  tm_bubbles(size = "binned_science",
#             col = "#034e7b", 
#             alpha = .8, 
#             title.size ="Avg. ACT Science Score", 
#             legend.size.is.portrait=TRUE,
#             sizes.legend.labels=c("<= 20.1","<= 20.6","<= 21.2","<= 22.6"), 
#             scale = 1.75) +
  #tm_layout(frame = FALSE, bg.color = "black") +
#  tm_legend(outside = TRUE, 
#            outside.position = "left", 
#            bg.color = "transparent", 
#            frame = FALSE)

###############################################################################

# Determine quantiles 
# quantile(my_map$cleaned_rural_mean_math, na.rm = TRUE)

# Set quantiles 
#my_map$binned_math <- as.numeric(my_map$cleaned_rural_mean_math)
#my_map$binned_math <- cut(my_map$cleaned_rural_mean_math,
#                             breaks = c(19.66250, 21.33818, 21.68000, 22.08706, 23.12000),
#                             labels = c(1, 2, 3, 4))
#my_map$binned_math <- as.numeric(my_map$binned_math)


#tm_shape(my_map) +
#  tm_fill("cleaned_rural_CountySpendingPerStudent", style = "quantile", n = 5, palette = "Greys", title = "Avg. Spending per Student") +
#  tm_bubbles(size = "binned_math",
#             col = "#034e7b",
#             alpha = .8, 
#             title.size ="Avg. ACT Math Score", 
#             legend.size.is.portrait=TRUE,
#             sizes.legend.labels=c("<= 21.3","<= 21.7","<= 22.1","<= 23.1"), 
#             scale = 1.75) +
  #tm_layout(frame = FALSE, bg.color = "black") +
#  tm_legend(outside = TRUE, 
#            outside.position = "left", 
#            bg.color = "transparent", 
#            frame = FALSE)


```





# Graduation Rate 
The below two maps provide the % of student who graduate from high school in four years for each county depicted as a proportional symbols. They are presented on a choropleth map showing the average amount of money spent per student. 

```{r, echo=FALSE}

# Define UI for the app
ui <- fluidPage(
  # Create two tabs for the maps
  navbarPage(
    "Maps",
    tabPanel("% 4yr Grad", 
             tmapOutput("map_grad")),
    tabPanel("% 5yr Grad",
             tmapOutput("map_grad5"))
  )
)

# Define server logic
server <- function(input, output) {
  # Render the science scores map
  output$map_grad <- renderTmap({
    my_map$binned_grad <- as.numeric(my_map$cleaned_rural_CountyGraduateRate4yr)
    my_map$binned_grad <- cut(my_map$cleaned_rural_CountyGraduateRate4yr,
                             breaks = c(81.9000, 91.1050, 94.0700, 96.0425, 100.0000),
                             labels = c(1, 2, 3, 4))
    my_map$binned_grad <- as.numeric(my_map$binned_grad)
    
    # Create the map
    tm_shape(my_map) +
      tm_fill("cleaned_rural_CountySpendingPerStudent", style = "quantile", n = 5, palette = "Greys", title = "Avg. Spending per Student") +
      tm_bubbles(size = "binned_grad",
                 col = "#6a51a3", 
                 alpha = .8, 
                 title.size ="4yr Graduation Rate", 
                 legend.size.is.portrait=TRUE,
                 sizes.legend.labels=c("<= 20.1","<= 20.6","<= 21.2","<= 22.6"), 
                 scale = 1.75) +
      tm_legend(outside = TRUE, 
                outside.position = "left", 
                bg.color = "transparent", 
                frame = FALSE)
  })
  
  # Render the math scores map
  output$map_grad5 <- renderTmap({
    # Set quantiles 
    my_map$binned_grad5 <- as.numeric(my_map$cleaned_rural_CountyGraduateRate5yr)
    my_map$binned_grad5 <- cut(my_map$cleaned_rural_CountyGraduateRate5yr,
                               breaks = c(88.08000, 93.45750, 95.26600, 97.39333, 100.0000),
                               labels = c(1, 2, 3, 4))
    my_map$binned_grad5 <- as.numeric(my_map$binned_grad5)

    
    # Create the map
    tm_shape(my_map) +
      tm_fill("cleaned_rural_CountySpendingPerStudent", style = "quantile", n = 5, palette = "Greys", title = "Avg. Spending per Student") +
      tm_bubbles(size = "binned_grad5",
                 col = "#6a51a3",
                 alpha = .8, 
                 title.size ="5yr Graduation Rate", 
                 legend.size.is.portrait=TRUE,
                 sizes.legend.labels=c("<= 93.5%","<= 95%","<= 97.5%","<= 100%"), 
                 scale = 1.75) +
      tm_legend(outside = TRUE, 
                outside.position = "left", 
                bg.color = "transparent", 
                frame = FALSE)
  })
}

# Run the app
shinyApp(ui = ui, server = server)

```


```{r, echo=FALSE}
# Determine quantiles 
# quantile(my_map$cleaned_rural_CountyGraduateRate4yr, na.rm = TRUE)

# Set quantiles 
#my_map$binned_grad <- as.numeric(my_map$cleaned_rural_CountyGraduateRate4yr)
#my_map$binned_grad <- cut(my_map$cleaned_rural_CountyGraduateRate4yr,
#                             breaks = c(81.9000, 91.1050, 94.0700, 96.0425, 100.0000),
#                             labels = c(1, 2, 3, 4))
#my_map$binned_grad <- as.numeric(my_map$binned_grad)


#tm_shape(my_map) +
#  tm_fill("cleaned_rural_CountySpendingPerStudent", style = "quantile", n = 5, palette = "Greys", title = "Avg. Spending per Student") +
#  tm_bubbles(size = "binned_grad",
#             col = "#6a51a3", 
#             n = 3, 
#             alpha = .8, 
#             title.size ="4yr Graduation Rate", 
#             legend.size.is.portrait=TRUE,
#             sizes.legend.labels=c("<= 91%","<= 94%","<= 96%","<= 100%"), 
#             scale = 1.75) +
  #tm_layout(frame = FALSE, bg.color = "black") +
#  tm_legend(outside = TRUE, 
#            outside.position = "right", 
#            bg.color = "transparent", 
#            frame = FALSE)

# Determine quantiles 
# quantile(my_map$cleaned_rural_CountyGraduateRate5yr, na.rm = TRUE)

# Set quantiles 
#my_map$binned_grad5 <- as.numeric(my_map$cleaned_rural_CountyGraduateRate5yr)
#my_map$binned_grad5 <- cut(my_map$cleaned_rural_CountyGraduateRate5yr,
#                             breaks = c(88.08000, 93.45750, 95.26600, 97.39333, 100.0000),
#                             labels = c(1, 2, 3, 4))
#my_map$binned_grad5 <- as.numeric(my_map$binned_grad5)


#tm_shape(my_map) +
#  tm_fill("cleaned_rural_CountySpendingPerStudent", style = "quantile", n = 5, palette = "Greys", title = "Avg. Spending per Student") +
#  tm_bubbles(size = "binned_grad5",
#             col = "#6a51a3", 
#             n = 3, 
#             alpha = .8, 
#             title.size ="5yr Graduation Rate", 
#             legend.size.is.portrait=TRUE,
#             sizes.legend.labels=c("<= 93.5%","<= 95%","<= 97.5%","<= 100%"), 
#             scale = 1.75) +
  #tm_layout(frame = FALSE, bg.color = "black") +
#  tm_legend(outside = TRUE, 
#            outside.position = "right", 
#            bg.color = "transparent", 
#            frame = FALSE)

```


# Levels of Education 
The below two maps provide the levels of education depicted as proportional symbols. They are presented on a choropleth map showing the average amount of money spent per student. 

```{r, echo=FALSE}
# Determine quantiles 
# quantile(my_map$cleaned_rural_Graduate.or.professional.degree.., na.rm = TRUE)


# Define UI for the app
ui <- fluidPage(
  # Create two tabs for the maps
  navbarPage(
    "Maps",
    tabPanel("% Bachelors Degree", 
             tmapOutput("binned_ba")),
    tabPanel("% Graduate Degree",
             tmapOutput("binned_phd"))
  )
)


# Define server logic
server <- function(input, output) {
  # Render the science scores map
  output$binned_phd <- renderTmap({
    my_map$binned_phd <- as.numeric(my_map$cleaned_rural_Graduate.or.professional.degree..)
    my_map$binned_phd <- cut(my_map$cleaned_rural_Graduate.or.professional.degree..,
                             breaks = c(2.7, 4.0, 4.8, 5.9, 23.7),
                             labels = c(1, 2, 3, 4))
    my_map$binned_phd <- as.numeric(my_map$binned_phd)
  
tm_shape(my_map) +
  tm_fill("cleaned_rural_CountySpendingPerStudent", style = "quantile", n = 5, palette = "Greys", title = "Avg. Spending per Student") +
  tm_bubbles(size = "binned_phd",
             col = "#02818a",
             n = 3, 
             alpha = .8, 
             title.size ="% Graduate Degree Holders", 
             legend.size.is.portrait=TRUE,
             sizes.legend.labels=c("<= 4%","<= 4.8%","<= 5.9%","<= 23.7%"), 
             scale = 1.75) +
  #tm_layout(frame = FALSE, bg.color = "black") +
  tm_legend(outside = TRUE, 
            outside.position = "left", 
            bg.color = "transparent", 
            frame = FALSE)

  })
  
# Determine quantiles 
# quantile(my_map$cleaned_rural_Bachelor.s.degree.., na.rm = TRUE)

  output$binned_ba <- renderTmap({
    my_map$binned_ba <- as.numeric(my_map$cleaned_rural_Bachelor.s.degree..)
    my_map$binned_ba <- cut(my_map$cleaned_rural_Bachelor.s.degree..,
                             breaks = c(7.40, 11.75, 13.50, 14.90, 32.10),
                             labels = c(1, 2, 3, 4))
    my_map$binned_ba <- as.numeric(my_map$binned_ba)


tm_shape(my_map) +
  tm_fill("cleaned_rural_CountySpendingPerStudent", style = "quantile", n = 5, palette = "Greys", title = "Avg. Spending per Student") +
  tm_bubbles(size = "binned_ba",
             col = "#02818a",
             alpha = .8, 
             title.size ="% Bachelor's Degree Holders", 
             legend.size.is.portrait=TRUE,
             sizes.legend.labels=c("<= 4%","<= 4.8%","<= 5.9%","<= 23.7%"), 
             scale = 1.75) +
  #tm_layout(frame = FALSE, bg.color = "black") +
  tm_legend(outside = TRUE, 
            outside.position = "left", 
            bg.color = "transparent", 
            frame = FALSE)

  })
}

# Run the app
shinyApp(ui = ui, server = server)
```


